Developing Address Clustering Heuristics for Account-Based Blockchain Networks: An Analysis based on a Specific Address Set

### 2023-08-26 paper.tex
Incorporate manual revision 1

\vspace{1em}
\onehalfspacing
\begin{center}
\section*{Abstract}
\end{center}
Given the negligible cost of creating blockchain addresses, the notion that one address equates to one user is flawed. Therefore, Blockchain Address Clustering emerges as a key tool for estimating the actual number of users. This paper examines heuristics to cluster addresses using Ethereum and Polygon PoS on-chain data, analyzing over 470,000 addresses collected from Decentraland over the course of nine months. We assess various existing heuristics, adapt them to our specific research context, and also introduce a novel heuristic. Our findings suggest that graph representation learning, particularly our variation of a node embedding algorithm, is the most effective approach for clustering a pre-defined address set. However, the method does not come without its drawbacks, which is why we propose possible extensions for improving clustering performance. \\

@@

Public permissionless blockchains such as Bitcoin \citep{nakamotoBitcoin2008} and Ethereum \citep{buterin2014ethereum} allow users to participate with multiple pseudonymous addresses, the creation of which is virtually cost-free. Contrary to popular belief, these blockchains are entirely transparent. All transactions are publicly observable and stored as part of the blockchain's history.
This circumstance has opened up a nascent scientific field dealing with entity identification within blockchain networks. Researchers try to cluster addresses controlled by the same entity by analyzing on-chain data and detecting usage patterns. The frequent lack of ground truth makes it challenging to evaluate different clustering methods. As a result, most methods are heuristic. \newline
On the one hand, identifying addresses that belong to the same real-world entity is beneficial as it allows for better evaluation of network properties concerning usage, distribution of wealth, and detecting fraudulent activities \citep{FV:17}. For instance, if users distribute their voting power across various addresses, they might manipulate an on-chain voting process that seems fair at the outset. \newline
Conversely, the lack of privacy is detrimental to most financial use cases. If someone links an address to an entity, they can observe all its transactions and activities. Using multiple addresses does not ensure privacy, as any connection between them can reveal that they belong to the same individual. \citep{nadler2023tornado} \newline
As a response, various privacy-enhancing tools and protocols have been proposed to obfuscate transaction tracing. 
Nonetheless, these protocols have yet to be widely adopted, and careless user behavior can undermine the privacy guarantees they offer. \newline
With these considerations in mind, privacy, or lack thereof, is and will remain an essential topic for blockchain development and research. %If someone obtains information that allows them to link an address to an entity, they may effectively observe that entity’s entire transaction history and associated activity. Even if the entity uses multiple addresses, any link between these addresses may expose the fact that they belong to the same person.

@@

Several address clustering heuristics in the UTXO setup have been proposed for Bitcoin and its derivatives \citep{Androulaki2013, Meiklejohn2013, Haslhofer2016, jourdan2018, kappos2022}. However, these methods are outside the scope of this work and will not be discussed further. \newline 

@@

\cite{wu2022tutela} extend on one of \cite{Beres2020} graph representation learning algorithms and apply it at scale. Further, they also propose a set of new clustering heuristics targeting Tornado Cash,  showing that careless user behavior can still reveal identity. Based on those heuristics, they developed an application to measure the anonymity of an Ethereum address. All of the aforementioned methods will be discussed in greater depth in Section 3. \newline
Broadening the scope of entity identification, \cite{victorlüders2019} study the largest Ethereum ERC-20 token networks from a graph perspective. Similarly, \cite{casalebrunet2021} apply network graph analysis to various non-fungible token (NFT) ecosystems. Both find that many networks follow a star or a hub-and-spoke pattern, similar to social network interaction graphs. Moreover, \cite{Payette2017} propose a segmentation of the Ethereum address space into four distinct behavior groups sharing similar attributes while applying k-means clustering. \newline

@@

\subsection{Contract Accounts and EOAs}
Account-based blockchains usually distinguish between externally owned accounts (EOAs) and contract accounts (smart contracts). \textit{EOAs} are created and controlled by private keys and can be used to hold the native protocol asset, send and receive transactions, and interact with contract accounts. \textit{Contract accounts} are controlled by the contract's code; their state can be modified through transactions sent to the contract, and they cannot initiate transactions. \citep{buterin2014ethereum} \newline Each account has a 20-byte address encoded in hexadecimal. For EOAs, this address is based on the last 20 bytes of the Keccak-256 hash of the ECDSA public key. For contract accounts, it is usually the last 20 bytes of the Keccak-256 hash of the RLP encoding of the sender address and account nonce. \citep{GW:14}

@@

\subsection{Decentraland}
\textit{Decentraland} is the first large-scale blockchain-based virtual world, with well-known companies like Tommy Hilfiger, Samsung, PepsiCo, Diesel, Adidas, and Netflix actively engaging in it \citep{metaverse-retailing2023}.\newline
Two main reasons make Decentraland an ideal platform for empirical research.
\textit{First}, its open architecture allows compiling snapshots of user activity, including the users' location within the metaverse and Web3 address. Using this capability, \cite{metaverse-retailing2023} captured snapshots of avatar locations over a nine-month period and provided the Web3 addresses collected for this thesis.\newline
\textit{Second}, users who connect their Web3 account can own various digital assets, such as monetary units, land parcels, and avatar collectibles (e.g., wearables, emotes, and names). Smart contracts on Ethereum and Polygon track and manage the ownership of these assets \citep{goldbergschaer2023}. This allows us to analyze the entire transaction history and derive information about the avatars' economic activity. \newline
Initially, Decentraland launched with two native tokens. \textit{LAND}, an NFT compliant with ERC-721, manages the ownership of land parcels in Decentraland. \textit{MANA}, a fungible token compliant with ERC-20, is the in-world currency used to purchase digital goods and services. The majority of purchases and trades are settled in MANA. \newline 
Moreover, the Decentraland core developers provide a smart-contract-based marketplace where users can buy or trade LAND or other in-game collectibles.
%The term metaverse refers to Neal Stephenson’s 1992 novel Snow Crash and describes an immersive virtual world that is populated by humans-as-avatars.

@@

In contrast, non-custodial crypto asset mixers (e.g., Tornado Cash) replace the trusted mixing party with a publicly verifiable smart contract. They rely on cryptographic schemes that allow anyone to prove and independently verify the validity of a withdrawal without disclosing the link to a specific deposit. This approach has the advantage that users do not have to share the identifying information with anyone, and there is no liquidity risk as the funds are locked. \citep{nadler2023tornado} \newline

@@

In theory, a larger anonymity set will provide a higher level of privacy since any withdrawal address could be linked to more deposit addresses. However, in practice, this does not always hold.
Some users reveal the link between their withdrawal and deposit addresses, which also affects the privacy of other depositors, for example, by withdrawing from the pool with the same address they deposited or directly sending funds between the two addresses. According to \cite{nadler2023tornado}, careless use of the protocol, testing transactions, or external incentive mechanisms might be reasons for this behavior. \newline

@@

\subsection{Transactions and Token Transfer Events}
The address subsets were used to collect transaction and token transfer event data from Ethereum and Polygon. Ethereum or Polygon data can be accessed directly through a node or an application programming interface (API) provider like Infura or Etherscan. The absence of account indexing in Ethereum or Polygon poses a challenge for retrieving all past transactions and token transfers of a specific address, as it requires scanning through all blocks and token transfer events emitted by designated token contracts. Fortunately, Etherscan offers an API Endpoint Module for ``Accounts`` that facilitates the retrieval of transactions and token transfer events for a given address. By using our network-specific address subsets, we were able to significantly decrease the number of required API calls. 
We gathered all transactions and token transfers up until block number \texttt{17,670,000} on Ethereum (July 11, 2023) and \texttt{44,990,000} on Polygon (July 12, 2023). In total, we collected over 30 million %30,689,978
token transfer events and more than 16 million %16,092,531
 normal transactions. While the amount of transactions is similar between the two networks, Polygon has roughly three times more transfer events. \newline The output was stored in separate CSV files and also imported into a MongoDB\footnote{\url{https://www.mongodb.com/}} database. The database contains distinct collections for transactions and transfers. For additional information on the data fields of each collection, see Appendix \ref{sec:database}.

@@

\subsection{ENS address pairs}
\label{sec:ens}
Similarly to \cite{Beres2020}, we use Ethereum Name Service (ENS) identifiers as ground truth information to evaluate some of the clustering methods.\newline
\textit{ENS} is a naming system that relies on the Ethereum smart contracts. Its main purpose is to map human-readable names, such as \texttt{`alice.eth`}, to machine-readable identifiers, mostly Ethereum addresses. The architecture of ENS comprises two main components: the registry and resolvers. \newline
The \textit{ENS registry} consists of a single smart contract\footnote{The registry contract is ERC-721 compliant, which means that .eth registrations can be transferred in the same way as other NFTs.} that maintains a list of all domains and subdomains, recording the ``owner`` and ``resolver`` for each. The registry allows the owner of a domain to make changes to that data. \citep{ENSdocs} %Domain owners may set the resolver for the domain.
\newline
\textit{Resolvers} are responsible for translating names into addresses. Any contract that adheres to the relevant standards may act as a resolver. The process of resolving a name involves two steps: \textit{First}, the registry must be queried to find out which resolver is responsible for the name, and \textit{second}, that resolver must be asked for the response to the query. \newline
In addition to regular resolution from name to address, ENS also supports ``reverse resolution``, allowing for a mapping from address back to a name or other metadata. Reverse resolution is accomplished via the %special
 domain \texttt{`addr.reverse`}% and the resolver function \texttt{name()}.
. This domain is owned by a special purpose registrar contract that allocates subdomains to the owner of the matching address - for instance, the address \texttt{`0x1234\dots`} may claim the name \texttt{`1234\dots.addr.reverse`}, and configure a resolver and records on it. The resolver in turn supports the \texttt{name()} function, which returns the ENS domain associated with that address.\newline
 Reverse resolution %(without a library)
 follows the same pattern as forward resolution: Get the resolver for \texttt{`1234....addr.reverse'} (where \texttt{`1234...`} is the address you want to reverse-resolve), and call the \texttt{name()} function on that resolver. However, ENS does not enforce the accuracy of reverse records. This means \texttt{`1234....addr.reverse`} may falsely claim that the name associated to their address is \texttt{`alice.eth`}.

Figure \ref{fig:ENS} illustrates the mechanism that allowed us to generate ground truth pairs in a simplified way: Initially, the domain name \texttt{`foo.eth`} resolves to address \texttt{`0xABC...`}. The owner has also set a reverse resolution for this address, indicated by the the edge from the address to the domain name. Next, the owner decides to change the resolved address to \texttt{`0xDEF...`}, again setting a reverse record.  As a domain can only point to one address at the time, the previous mapping from \texttt{`foo.eth`} to the old address is deleted. However, \texttt{`ABC....addr.reverse`} still points to \texttt{`foo.eth`} and when we query the name corresponding to \texttt{`0xABC...`} and \texttt{`0xDEF...`}, both will return \texttt{`foo.eth`}.
If this special case occurs, we assume that both addresses belong to the same entity.
%We did not observe cases where someone set a reverse record to a completely unrelated address (e.g. alice.eth)

@@

		\end{tikzpicture}
		\caption{Changing the record to another address does not change the old reverse resolution (Simplified Schematic Representation)}
		\label{fig:ENS}
	\end{center}
\end{figure*}

We applied the following methodology to find ground truth pairs: \textit{First}, we gathered all addresses that interacted with ENS to reduce the amount of name look-ups. \textit{Second}, we iterated over each address and checked if it reverse-resolved to an ENS domain\footnote{This was done via the ENS module within web3.py \url{https://web3py.readthedocs.io/en/stable/ens_overview.html}, which provides an interface to look up domains and addresses.}. If so, we logged the ENS domain and address in a CSV file. \textit{Third}, using our new list, we identified if two addresses referenced the same ENS domain. If matched, we recorded both addresses alongside the ENS domain. \newline
From this, we identified 11,440 reverse records and 40 pairs of addresses pointing to the same ENS name, which we utilized as our ground truth to evaluate various clustering heuristics. Although this sample is relatively small, it allows us to assess and draw conclusions about the effectiveness of the methods.

@@

\subsection{Self-authorization} 
	All three token standards require functions to allow another address to transfer tokens on behalf of the actual owner. For the ERC-20 standard, this is achieved through the \texttt{`approve`} function, accepting the spender's address and the token amount as parameters\footnote{\texttt{approve(address spender, uint256 amount)}}. The ERC-721 has two functions dedicated to approvals: One for authorizing the transfer of a specific NFT\footnote{\texttt{approve(address to, uint256 tokenId)}} and one to permit transferring all of the owner's NFTs within the collection\footnote{\texttt{setApprovalForAll(address operator, bool approved)}}. The latter function also appears in the ERC-1155 standard. The approval functionality of these tokens is primarily used in connection with smart contracts (e.g., Decentralized Exchanges or NFT Marketplaces), but can also be applied for regular EOAs. \newline
The ``self authorization`` heuristic, proposed by \cite{FV:17}, suggests that users might authorize another address they own, possibly for risk distribution across several addresses or for testing purposes. \newline
Since all the addresses' transactions are available in our data set, we can easily employ this heuristic.

\subsection{Deposit address reuse}
The concept of deposit address reuse was first introduced and systematically utilized by \cite{FV:17} and subsequently adoped by \cite{wu2022tutela}. This heuristic relies on the prevalent practice of crypto exchanges generating unique deposit addresses for each user. These addresses then forward the funds to an exchange's main address. It is highly likely that multiple addresses sending funds to the same deposit address are controlled by the same entity. The key challenge of this approach lies in identifying these deposit addresses, which could either be EOAs or smart contracts. \citep{FV:17}\newline
Deposit addresses share the characteristic that they forward the received funds to a major exchange account. Two parameters are essential for deposit address detection: the discrepancy between the amount received and forwarded (since the exchange must pay for gas fees), and the time lag between receiving and forwarding funds.
However, our data set only includes the direct transactions or token transfers associated with the addresses. The transactions between deposit and main exchange address are absent. Consequently, this clustering heuristic is not applicable %with our data
 and it would require significantly more effort for gathering the required data.

\subsection{Airdrop multi participation}
Airdrops are popular token distribution mechanisms, and recipients are mostly chosen based on past protocol activity. Probably the most well known example is Uniswap's UNI airdrop, where each address that interacted with the protocol until a specified cutoff date received a fixed amount of 400 UNI tokens\footnote{\url{https://blog.uniswap.org/uni}}. Especially in the early days of airdrops, these distribution mechanisms were not Sybil-resistant, leading to people creating multiple addresses in anticipation of an airdrop. However, \cite{FV:17} argues that managing the tokens on all of these addresses is impractical, which is why they are often aggregated into one address. He finds around 500 same-source, fixed amount token distribution events with at least 1,000 recipients while also taking into account the block difference between the individual aidrop token transfers. He defines multi-participation when two airdrop recipients forward their tokens to a single address. Since we do not have a whole view of the airdrop distribution events, we are not able to identify them within our data. One possible approach would include manual selection of known airdrops. However, this process takes very long and is not guaranteed to cluster a lot of addresses, since most new airdrop events were carried out with some kind of sybil-resistance. Nevertheless, to test this approach we searched for addresses that forwarded the UNI airdrop to another address. We found eight such transfers, but none involved two addresses sending funds to a third address.

@@

According to \cite{Beres2020}, \textit{node embedding} methods form a class of graph representation learning\footnote{Sometimes also referred to as ``network representation learning``} methods that map graph nodes to vectors in a low-dimensional vector space. These methods aim to represent vertices with similar neighborhood structure by vectors that are close in the vector space. \newline

@@

The foundation for these node-embedding methods is an undirected graph where nodes are composed of distinct addresses, and an edge is placed between two nodes if there is a transaction between them. \newline
To construct the network graph, we use the intra-set asset transfers described in section \ref{sec:intra-set}. This approach differs from the methodologies adopted by \cite{Beres2020} and \cite{wu2022tutela}. On the one hand, \cite{Beres2020} collected all transactions and token transfer events of their address set but did not refine this data to only include intra-set asset transfers. On the other hand, \cite{wu2022tutela} do not operate on a specific address set but use all Ethereum transactions to build the network graph, ignoring token transfer events. After comparing these two methodologies with ours, we determined that intra-set asset transfers were the most effective for clustering a pre-defined set of addresses. %using node-embedding methods.

@@

Using the network graph, node embedding methods seek to learn a function that projects a node to a $d$-dimensional vector representation (also called feature vector). This is a way of simplifying the graph information by associating each node with a point in the Euclidean space. Various node embedding methods have been developed and applied to different domains. \citep{rozemberczki2020difftovec} \newline
The intuition behind node-embedding approaches is that addresses which interact with the same set of addresses within the network graph should be close in Euclidean distance. \cite{Beres2020} cites two main reasons for this. Firstly, users with multiple accounts often interact with the same addresses or services from most of them. Secondly, when users move funds between their personal addresses, they may unintentionally reveal their address clusters. \newline
\cite{karateclub} provide a variety of node embedding methods as a Python library\footnote{\url{https://github.com/benedekrozemberczki/karateclub}}. They categorize the methods into neighborhood (proximity) preserving and structural. \newline
We focus on the three methods that performed best in \cite{Beres2020} experiments: ``\textit{Diff2Vec}`` \citep{rozemberczki2020difftovec}, ``\textit{Role2Vec}`` \citep{ahmed2018roletovec}, and ``\textit{Deepwalk}`` \citep{perozzi2014}. Diff2Vec and DeepWalk are neighborhood-preserving methods, while Role2Vec is a structural node embedding method. Although all methods utilize random walks to capture graph information, their approaches to leveraging these random walks for constructing Euclidean feature vectors differ. A detailed explanation of each method is beyond the scope of this thesis. The interested reader may consult the original publications. For a concise overview of Diff2Vec, please refer to \cite{wu2022tutela}. \newline
With our intra-set asset transfer data (\ref{sec:intra-set}) and ENS address pairs (\ref{sec:ens}), we were able to apply and evaluate the node embedding methods described above. Using the library described by \cite{karateclub} requires certain preprocessing steps for the network graph. Transactions must be treated as undirected edges and loops (self-transactions), and multi-edges must be removed. Moreover, nodes outside the largest connected component are excluded. The resulting network graph includes 51,566 nodes and 249,302 edges. \newline
It is worth noting that unlike \cite{wu2022tutela}, we decided not to include edge weights. For instance, if Alice sent Bob Ether three times and Bob sent Alice Ether two times, the edge weight would correspond to five. This choice facilitated the application of \cite{rozemberczki2020difftovec}'s library and eliminated the need to create a custom method. Based on our network graph, we generate 128-dimensional feature vectors for each address and method. \newline

>>>>

\vspace{1em}
\onehalfspacing
\begin{center}
\section*{Abstract}
\end{center}
Given the negligible cost of creating blockchain addresses, the notion that one address equates to one user is flawed. Therefore, Blockchain Address Clustering emerges as a key tool for estimating the actual number of users. This paper examines heuristics to cluster addresses using Ethereum and Polygon PoS data, analyzing over 470,000 addresses collected from Decentraland over the course of nine months. We assess various existing heuristics, adapt them to our specific research context, and also introduce a novel heuristic. Our findings suggest that graph representation learning, particularly our variation of a node embedding algorithm, is the most effective approach for clustering a predefined address set. However, the method does not come without its drawbacks, which is why we propose possible extensions for improving clustering performance. \\

@@

Public permissionless blockchains such as Bitcoin \citep{nakamotoBitcoin2008} and Ethereum \citep{buterin2014ethereum} allow users to participate with multiple pseudonymous addresses, the creation of which is virtually cost-free. Contrary to popular belief, these blockchains are entirely transparent. All transactions are publicly observable and stored as part of the blockchain's history.
This circumstance has opened up a nascent scientific field dealing with entity identification within blockchains. Researchers try to cluster addresses controlled by the same entity by analyzing on-chain data and detecting usage patterns. The frequent lack of ground truth makes it challenging to evaluate different clustering methods. As a result, most methods are heuristic. \newline
On the one hand, identifying addresses that belong to the same real-world entity is beneficial as it allows for better evaluation of network properties concerning usage, distribution of wealth, and detecting fraudulent activities \citep{FV:17}. For instance, if users distribute their voting power across various addresses, they might manipulate an on-chain voting process that seems fair at the outset. \newline
Conversely, the lack of privacy is detrimental to most financial use cases. If someone links an address to an entity, they can observe all its transactions and associated activities. Using multiple addresses does not ensure privacy, as any connection between them can reveal that they belong to the same individual. \citep{nadler2023tornado} \newline
As a response, various privacy-enhancing tools and protocols have been proposed to obfuscate transaction tracing. 
Nonetheless, these protocols have yet to be widely adopted, and careless user behavior can undermine the privacy guarantees they offer. \newline
With these considerations in mind, privacy, or lack thereof, is and will remain an essential topic for blockchain development and research. %If someone obtains information that allows them to link an address to an entity, they may effectively observe that entity’s entire transaction history and associated activity. Even if the entity uses multiple addresses, any link between these addresses may expose the fact that they belong to the same person.

@@

Several address clustering heuristics in the UTXO setup have been proposed for Bitcoin and its derivatives \citep{Androulaki2013, Meiklejohn2013, Haslhofer2016, jourdan2018, kappos2022}. However, these methods are outside the scope of this thesis and will not be discussed further. \newline 

@@

\cite{wu2022tutela} extend on one of \cite{Beres2020} graph representation learning algorithms and apply it at scale. Further, they propose a set of new clustering heuristics targeting Tornado Cash,  showing that careless user behavior can still reveal identity. Based on those heuristics, they developed an application to measure the anonymity of an Ethereum address. All of the aforementioned methods will be discussed in greater depth in Section 3. \newline
Broadening the scope of entity identification, \cite{victorlüders2019} study the largest fungible token networks on Ethereum from a graph perspective. Similarly, \cite{casalebrunet2021} apply network graph analysis to various non-fungible token (NFT) ecosystems. Both find that many networks follow either a star or a hub-and-spoke pattern, similar to social network interaction graphs. Moreover, \cite{Payette2017} propose a segmentation of the Ethereum address space into four distinct behavior groups sharing similar attributes, while applying k-means clustering. \newline

@@

\subsection{Contract Accounts and EOAs}
Account-based blockchains usually distinguish between externally owned accounts (EOAs) and contract accounts (smart contracts). \textit{EOAs} are created and controlled by private keys and can be used to hold the native protocol asset, send and receive transactions, and interact with contract accounts. \textit{Contract accounts} are controlled by the contract's code; their state can be modified through transactions sent to the contract, and they cannot initiate transactions. \citep{buterin2014ethereum} \newline Each account has a 20-byte address encoded in hexadecimal. For EOAs, this address is based on the last 20 bytes of the Keccak-256 hash of the ECDSA public key. For contract accounts, it is usually the last 20 bytes of the Keccak-256 hash derived from the sender's address and account nonce. \citep{GW:14}

@@

\subsection{Decentraland}
\textit{Decentraland} is the first large-scale blockchain-based virtual world, with well-known companies like Tommy Hilfiger, Samsung, PepsiCo, Diesel, Adidas, and Netflix actively engaging in it \citep{metaverse-retailing2023}.\newline
Two main reasons make Decentraland an ideal platform for empirical research.
\textit{First}, its open architecture allows compiling snapshots of user activity, including the users' location within the metaverse and Web3 address. Using this capability, \cite{metaverse-retailing2023} captured snapshots of avatar locations over a nine-month period and provided the collected addresses for this thesis.\newline
\textit{Second}, users who connect their Web3 account can own various digital assets, such as monetary units, land parcels, and avatar collectibles (e.g., wearables, emotes, and names). Smart contracts on Ethereum and Polygon track and manage the ownership of these assets \citep{goldbergschaer2023}. This allows us to analyze the entire transaction history and derive information about the avatars' economic activity. \newline
Initially, Decentraland launched with two native tokens. \textit{LAND}, an NFT compliant with ERC-721, manages the ownership of land parcels in Decentraland. \textit{MANA}, a fungible token compliant with ERC-20, is the in-world currency used to purchase digital goods and services. The majority of purchases and trades are settled in MANA. \newline 
Moreover, the Decentraland core developers provide a smart-contract-based marketplace where users can buy or trade LAND or other in-game collectibles.

@@

In contrast, non-custodial crypto asset mixers, such as Tornado Cash, replace the trusted mixing party with a publicly verifiable smart contract. They rely on cryptographic schemes that allow anyone to prove and independently verify the validity of a withdrawal without disclosing the link to a specific deposit. This approach has the advantage that users do not have to share the identifying information with anyone, and there is no liquidity risk as the funds are locked. \citep{nadler2023tornado} \newline

@@

In theory, a larger anonymity set will provide a higher level of privacy since any withdrawal address could be linked to more deposit addresses. However, in practice, this principle does not always hold.
Some users reveal the link between their withdrawal and deposit addresses, for example, by withdrawing from the pool with the same address they deposited or directly sending funds between the two addresses. This behavior also affects the privacy of other depositors. According to \cite{nadler2023tornado}, careless use of the protocol, testing transactions, or external incentive mechanisms might be reasons for this behavior. \newline

@@

\subsection{Transactions and Token Transfer Events}
The address subsets were used to collect transaction and token transfer event data from Ethereum and Polygon. Ethereum or Polygon data can be accessed directly through a node or an application programming interface (API) provider like Infura or Etherscan. The absence of account indexing in Ethereum or Polygon poses a challenge for retrieving all past transactions and token transfers of a specific address, as it requires scanning through all blocks and token transfer events emitted by designated token contracts. Fortunately, Etherscan offers an API Endpoint for ``Accounts`` that facilitates the retrieval of transactions and token transfer events for a given address. By using our network-specific address subsets, we were able to significantly decrease the number of required API calls. 
We gathered all transactions and token transfers up until block number \texttt{17,670,000} on Ethereum (July 11, 2023) and \texttt{44,990,000} on Polygon (July 12, 2023). In total, we collected over 30 million %30,689,978
token transfer events and more than 16 million %16,092,531
 normal transactions. While the amount of transactions is similar between the two networks, Polygon has roughly three times more transfer events. \newline The output was stored in separate CSV files and also imported into a MongoDB\footnote{\url{https://www.mongodb.com/}} database. The database contains distinct collections for transactions and transfers. For additional information on the data fields of each collection, see Appendix \ref{sec:database}.

@@

\subsection{ENS address pairs}
\label{sec:ens}
Similarly to \cite{Beres2020}, we use Ethereum Name Service (ENS) identifiers as ground truth information to evaluate clustering methods.\newline
\textit{ENS} is a naming system that relies on the Ethereum smart contracts. Its main purpose is to map human-readable names, such as \texttt{`alice.eth`}, to machine-readable identifiers, mostly Ethereum addresses. The architecture of ENS comprises two main components: the registry and resolvers. \newline
The \textit{ENS registry} consists of a single smart contract\footnote{The registry contract is ERC-721 compliant, which means that .eth registrations can be transferred in the same way as other NFTs.} that maintains a list of all domains and subdomains, recording the ``owner`` and ``resolver`` for each. The registry allows the owner of a domain to make changes to that data. 
\textit{Resolvers} are responsible for translating names into addresses. Any contract that adheres to the relevant standards may act as a resolver. The process of resolving a name involves two steps: \textit{First}, the registry must be queried to find out which resolver is responsible for the name, and \textit{second}, that resolver must be asked for the response to the query. \citep{ENSdocs} \newline
In addition to regular resolution from name to address, ENS also supports ``reverse resolution``, allowing for a mapping from address back to a name or other metadata. Reverse resolution is accomplished via the %special
 domain \texttt{`addr.reverse`}% and the resolver function \texttt{name()}.
. This domain is owned by a special purpose registrar contract that allocates subdomains to the owner of the matching address - for instance, the address \texttt{`0x1234\dots`} may claim the name \texttt{`1234\dots.addr.reverse`}, and configure a resolver and records on it. The resolver in turn supports the \texttt{name()} function, which returns the ENS domain associated with that address.\newline
 Reverse resolution %(without a library)
 follows the same pattern as forward resolution: Get the resolver for \texttt{`1234....addr.reverse'} (where \texttt{`1234...`} is the address you want to reverse-resolve), and call the \texttt{name()} function on that resolver. However, ENS does not enforce the accuracy of reverse records. This means \texttt{`1234....addr.reverse`} may falsely claim that the name associated to their address is \texttt{`alice.eth`}.

Figure \ref{fig:ENS} illustrates the mechanism that allowed us to generate ground truth pairs in a simplified way: Initially, the domain name \texttt{`foo.eth`} resolves to address \texttt{`0xABC...`}. The owner has also set a reverse resolution for this address, indicated by the the edge from the address to the domain name. Next, the owner decides to change the resolved address to \texttt{`0xDEF...`}, again setting a reverse record.  As a domain can only point to one address at the time, the previous mapping from \texttt{`foo.eth`} to the old address is deleted. However, \texttt{`ABC....addr.reverse`} still points to \texttt{`foo.eth`} and when we query the name corresponding to \texttt{`0xABC...`} and \texttt{`0xDEF...`}, both will return \texttt{`foo.eth`}.
If this special case occurs, we assume that both addresses belong to the same entity.

@@

		\end{tikzpicture}
		\caption{Simplified Schematic Representation of a Record Change to Another Address}
		\label{fig:ENS}
	\end{center}
\end{figure*}

We applied the following methodology to find ground truth pairs: \textit{First}, we gathered all addresses that interacted with ENS to reduce the amount of name look-ups. \textit{Second}, we iterated over all addresses and checked if they reverse-resolved to an ENS domain\footnote{This was done via the ENS module within web3.py \url{https://web3py.readthedocs.io/en/stable/ens_overview.html}, which provides an interface to look up domains and addresses.}. If so, we logged the ENS domain and address in a CSV file. \textit{Third}, using our new list, we identified if two addresses referenced the same ENS domain. If matched, we recorded both addresses alongside the ENS domain. \newline
From this, we identified 11,440 reverse records and 40 pairs of addresses pointing to the same ENS name, which we utilized as our ground truth to evaluate various clustering heuristics. Although this sample is relatively small, it allows us to assess and draw conclusions about the effectiveness of the methods.

@@

\subsection{Self-authorization} 
	All three token standards require functions to allow another address to transfer tokens on behalf of the actual owner. For the ERC-20 standard, this is achieved through the \texttt{`approve`} function, accepting the spender's address and the token amount as parameters\footnote{\texttt{approve(address spender, uint256 amount)}}. The ERC-721 has two functions dedicated to approvals: One for authorizing the transfer of a specific NFT\footnote{\texttt{approve(address to, uint256 tokenId)}} and one to permit transferring all of the owner's NFTs within the collection\footnote{\texttt{setApprovalForAll(address operator, bool approved)}}. The latter function also appears in the ERC-1155 standard. The approval functionality of these tokens is primarily used in connection with smart contracts (e.g., Decentralized Exchanges or NFT Marketplaces), but can also be applied for regular EOAs. \newline
The ``self authorization`` heuristic, proposed by \cite{FV:17}, suggests that users might authorize another address they own, possibly for risk distribution across several addresses or for testing purposes. Since all the addresses' transactions are available in our data set, we can easily employ this heuristic.

\subsection{Deposit address reuse}
The concept of deposit address reuse was first introduced and systematically utilized by \cite{FV:17} and subsequently adoped by \cite{wu2022tutela}. This heuristic relies on the prevalent practice of crypto exchanges generating unique deposit addresses for each user. These addresses then forward the funds to an exchange's main address. It is highly likely that multiple addresses sending funds to the same deposit address are controlled by the same entity. The key challenge of this approach lies in identifying these deposit addresses, which could either be EOAs or smart contracts. \citep{FV:17}\newline
Deposit addresses share the characteristic that they forward the received funds to a major exchange account. Two parameters are essential for deposit address detection: the discrepancy between the amount received and forwarded (since the exchange must pay for gas fees), and the time lag between receiving and forwarding funds.
However, our data set only includes the transactions or token transfers associated with the addresses. The transactions between deposit and exchange address are absent. Consequently, this clustering heuristic is not applicable and would require significantly more effort for gathering the required data.

\subsection{Airdrop multi participation}
Airdrops are popular token distribution mechanisms, and recipients are mostly chosen based on past protocol activity. Probably the most well known example is Uniswap's UNI airdrop, where each address that interacted with the protocol until a specified cutoff date received a fixed amount of 400 UNI tokens\footnote{\url{https://blog.uniswap.org/uni}}. Especially in the early days of airdrops, these distribution mechanisms were not Sybil-resistant, leading to people creating multiple addresses in anticipation of an airdrop. However, \cite{FV:17} argues that managing the tokens on all of these addresses is impractical, which is why they are often aggregated into one address. He found around 500 same-source, fixed amount token distribution events with at least 1,000 recipients while also taking into account the block difference between the individual aidrop token transfers. He defined multi-participation when two airdrop recipients forward their tokens to a single address. Since we do not have a whole view of the airdrop distribution events, we are not able to identify them within our data. One possible approach would include manual selection of known airdrops. However, this process takes a long time and is not guaranteed to cluster many addresses, since more recent airdrop events were mostly carried out with some kind of sybil-resistance. Nevertheless, to test this approach we searched for addresses that forwarded the UNI airdrop to another address. We found eight such transfers, but none involved two addresses sending funds to a third address.

@@

According to \cite{Beres2020}, \textit{node embedding} algorithms form a class of graph representation learning\footnote{Sometimes also referred to as ``network representation learning``} methods that map graph nodes to vectors in a low-dimensional vector space. These methods aim to represent vertices with similar neighborhood structure by vectors that are close in the vector space. \newline

@@

The foundation for these node-embedding algorithms is an undirected graph where nodes are composed of distinct addresses, and an edge is placed between two nodes if there is a transaction between them. \newline
To construct the network graph, we use the intra-set asset transfers described in section \ref{sec:intra-set}. Our approach differs from the methodologies adopted by \cite{Beres2020} and \cite{wu2022tutela}. On the one hand, \cite{Beres2020} collected all transactions and token transfer events of their address set but did not refine this data to only include intra-set asset transfers. On the other hand, \cite{wu2022tutela} do not operate on a specific address set but use all Ethereum transactions to build the network graph, ignoring token transfer events. After comparing these two methodologies with ours, we determined that intra-set asset transfers were the most effective for clustering a predefined set of addresses. %using node-embedding methods.

@@

Using the network graph, node embedding algorithms seek to learn a function that projects a node to a $d$-dimensional vector representation (also called feature vector). This is a way of simplifying the graph information by associating each node with a point in the Euclidean space. Various node embedding algorithms have been developed and applied to different domains. \citep{rozemberczki2020difftovec} \newline
The intuition behind node embeddings is that addresses which interact with the same set of addresses within the network graph should be close in Euclidean distance. \cite{Beres2020} cites two main reasons for this. \textit{First}, users with multiple accounts often interact with the same addresses or services from most of them. \textit{Second}, when users move funds between their personal addresses, they may unintentionally reveal their address clusters. \newline
\cite{karateclub} provide a variety of node embedding algorithms as a Python library\footnote{\url{https://github.com/benedekrozemberczki/karateclub}}. They categorize the methods into neighborhood (proximity) preserving and structural. \newline
We focus on the three methods that performed best in \cite{Beres2020} experiments: ``\textit{Diff2Vec}`` \citep{rozemberczki2020difftovec}, ``\textit{Role2Vec}`` \citep{ahmed2018roletovec}, and ``\textit{DeepWalk}`` \citep{perozzi2014}. Diff2Vec and DeepWalk are neighborhood-preserving methods, while Role2Vec is a structural node embedding method. Although all methods utilize random walks to capture graph information, their approaches to leveraging these random walks for constructing Euclidean feature vectors differ. A detailed explanation of each algorithm is beyond the scope of this thesis. The interested reader may consult the original publications. For a concise overview of Diff2Vec, please refer to \cite{wu2022tutela}. \newline
With our intra-set asset transfer data (\ref{sec:intra-set}) and ENS address pairs (\ref{sec:ens}), we were able to apply and evaluate the node embedding methods described above. To use the library described by \cite{karateclub}, certain preprocessing steps for the network graph are required. Transactions must be treated as undirected edges, loops (self-transactions) and multi-edges must be removed. Moreover, nodes outside the largest connected component are excluded. The resulting network graph includes 51,566 nodes and 249,302 edges. \newline
